\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{References}
\begin{thebibliography}{99}

\bibitem{Diestel2012}
    R. Diestel
    \emph{Graph Theory -- Graduate Texts in Mathematics}.
    Springer-Diestel,
    4th Edition, corrected reprint,
    2012.

\bibitem{AsratianDenleyHaggkvist}
  A.S. Asratian, T.M.J. Denley, R. Häggkvist,
  \emph{Bipartite Graphs and their Applications}.
  Cambridge Tracts in Mathematics \textbf{131},
  Cambridge University Press,
  1998.

\bibitem{Scheinerman2012}
  E.R. Scheinerman, 
  \emph{Mathematics: A Discrete Introduction}.
  3rd Edition,
  Cengage Learning,
  2012.

\bibitem{ChartrandZhang2008}
  G. Chartrand, P. Zhang,
  \emph{Chromatic Graph Theory}.
  Discrete Mathematics And Its Applications \textbf{53},
  CRC Press,
  2008.

\bibitem{WassermanFaust1994}
  S. Wasserman, K. Faust,
  \emph{Social Network Analysis: Methods and Applications}.
  Structural Analysis in the Social Sciences \textbf{8},
  Cambridge University Press,
  1994.

\bibitem{Niedermeier2006}
  R. Niedermeier,
  \emph{Invitation to Fixed Parameter Algorithms}.
  Oxford Lecture Series in Mathematics and Its Applications,
  Oxford,
  2006.

\bibitem{Trudeau1993}
  R.J. Trudeau,
  \emph{Introduction to Graph Theory},
  New York: Dover Pub,
  1993.

\bibitem{Harris2000}
  J.M. Harris
  \emph{Combinatorics and Graph Theory}.
  Springer-Verlag,
  2000.
  
\bibitem{HollandLeinhardt1971}
  P.W. Holland, S. Leinhardt
  \emph{Transitivity in structural models of small groups}.
  Comparative Group Studies \textbf{2} (2), pp 107-24,
  1971.

\bibitem{WattsStrogatz1998}
  D.J. Watts, S. Strogatz
  \emph{Collective dynamics of \textquote{small-world} networks}.
  Nature \textbf{393} (6684), pp 440–42,
  1998.
  
\bibitem{Freeman1977}
  L.C. Freeman
  \emph{A set of measures of centrality based on betweenness}.
  Sociometry \textbf{40} (1), pp 35-41,
  1977.

\bibitem{LucePerry1949}
  R.D. Luce, A.D. Perry
  \emph{A method of matrix analysis of group structure}.
  Psychometrika \textbf{14} (2), pp 95-116,
  1949.

\bibitem{WassermanFaust1994}
  S. Wasserman, K. Faust
  \emph{Social Network Analysis: Methods and Applications}.
  Cambridge University Press,
  1994.

\bibitem{AmaralOttino2004}
	L.A.N. Amaral, J.M. Ottino,
	\emph{Complex networks. Augmenting the framework for the study of complex systems}.
	The European Physical Journal B \textbf{38}, pp 147-62,
	2004.
	
\bibitem{Davis2012}
  Mark Davis,
  \emph{Unicode Text Segmentation}.
  Unicode Standard Annex \textbf{\#29}, http://www.unicode.org/reports/tr29/,
  2012.
  
\bibitem{Turing1950}
  A. Turing,
  \emph{Computing Machinery and Intelligence}.
  Mind \textbf{LIX},
  1950.

\bibitem{ManningSchutze1999}
  C.D. Manning, H. Schütze,
  \emph{Foundations of Statistical Natural Language Processing}.
  MIT Press,
  1999.

\bibitem{Auerbach1913}
  F. Auerbach,
  \emph{Das Gesetz der Bevölkerungskonzentration}.
  Petermanns Geogr Mitt \textbf{59},
  1913.

\bibitem{Eeckhout2004}
  J. Eeckhout,
  \emph{Gibrat's law for (All) Cities}.
  American Economic Review \textbf{94} (5), pp 1429-51,
  2004.

\bibitem{AdamicHuberman2002}
  L.A. Adamic, B.A. Huberman,
  \emph{Zipf’s law and the Internet}.
  Glottometrics \textbf{3}, pp 143-50,
  2002.

\bibitem{Brillouin2004}
  Léon Brillouin,
  \emph{Science and Information Theory}.
  Academic Press, New York,
  English translation of 1988 reprint,
  2004.

\bibitem{WentianLi1992}
  Wentian Li,
  \emph{Random Texts Exhibit Zipf's-Law-Like Word Frequency Distribution}.
  IEEE Transactions on Information Theory \textbf{38} (6), pp 1842-45,
  1992.

\bibitem{Neumann2013}
  P.G. Neumann,
  \emph{Statistical metalinguistics and Zipf/Pareto/Mandelbrot}.
  SRI International Computer Science Laboratory,
  accessed and archived 29 Aug 2013 (http://web.archive.org/web/20130829073335/http://www.csl.sri.com/users/neumann/).

\bibitem{Belevitch1959}
  V. Belevitch,
  \emph{On the statistical laws of linguistic distributions}.
  Annales de la Société Scientifique de Bruxelles. I \textbf{73}, pp 310-26,
  1959.

\bibitem{Zipf1949}
  G.K. Zipf,
  \emph{Human Behavior and the Principle of Least Effort}.
  Cambridge, Massachusetts: Addison-Wesley,
  1949.

\bibitem{CanchoSole2003}
  R.F. i Cancho, R.V. Sole,
  \emph{Least effort and the origins of scaling in human language}.
  Proceedings of the National Academy of Sciences of the United States of America \textbf{100} (3), pp 788-91,
  2003.

\bibitem{OnnelaSaramakiBarabasi2007}
  J.P. Onnela, J. Saramaki, J. Hyvonen, G. Szabo, D. Lazer, K. Kaski, J. Kertesz, A.L. Barabasi,
  \emph{Structure and tie strengths in mobile communication networks},
  Proceedings of the National Academy of Sciences \textbf{104} (18), pp 7332-36,
  2007.

\bibitem{ChoromanskiMatuszakMiekisz2013}
  K. Choromański, M. Matuszak, J. Miękisz,
  \emph{Scale-Free Graph with Preferential Attachment and Evolving Internal Vertex Structure}.
  Journal of Statistical Physics \textbf{151} (6), pp 1175-83,
  2013.

\bibitem{ClausetShaliziNewman2007}
  A. Clauset, C.R. Shalizi, M.E.J Newman,
  \emph{Power-law distributions in empirical data}.
  SIAM Review \textbf{51}, pp 661-703,
  2009.

\bibitem{AielloChungLu2000}
  W. Aiello, F. Chung, L. Lu,
  \emph{A random graph model for massive graphs},
  in Proceedings of the 32nd Annual ACM Symposium on Theory of Computing, pp 171-80,
  Association of Computing Machinery, New York,
  2000.

\bibitem{AdamicLukosePuniyaniHuberman2001}
  L.A. Adamic, R.M. Lukose, A.R. Puniyani, B.A. Huberman,
  \emph{Search in power-law networks},
  Physical Review E \textbf{64}, 046135,
  2001

\bibitem{CohenErezAvrahamHavlin2000}
  R. Cohen, K. Erez, D. ben-Avraham, S. Havlin,
  \emph{Resilience of the Internet to random breakdowns}.
  Physical Review Letters \textbf{85}, 4626,
  2000.

\bibitem{DorogovtsevSamukhin2002}
  S.N. Dorogovtsev, A.N. Samukhin,
  \emph{Mesoscopics and fluctuations in networks}.
  Physical Review E \textbf{67}, 037103,
  2003.

\bibitem{DorogovtsevMendesSamukhin2001}
  S.N. Dorogovtsev, J.F.F. Mendes, A.N. Samukhin,
  \emph{Size-dependent degree distribution of a scale-free growing network}.
  Physical Review E \textbf{6Physical Review Letters3} (6), 062101,
  2001.

\bibitem{BarabasiAlbert1999}
  A.L. Barabási, R. AlPhysical Review Ebert,
  \emph{Emergence of scaling in random networks}.
  Science \textbf{286} (5439), pp 509-12,
  1999.

\bibitem{BarabasiAlbert2002}
  A.L. Barabási, R. Albert,
  \emph{Statistical mechanics of complex networPhysycal Review Lettersks}.
  Reviews of Modern Physycs \textbf{74} (1), pp 47-97,
  2002.
  
\bibitem{MoreiraAndradeAmaral2002}
  A.A. Moreira, J.S. Andrade Jr., L.A.N. Amaral,
  \emph{Extremum statistics in scale-free network models}.
  Physical Review Letters \textbf{89} (26), 268703,
  2002.
  
\bibitem{Egghe2007}
  L. Egghe,
  \emph{Untangling Herdan's law and Heaps' law: Mathematical and informetric arguments}.
  Journal of the American Society for Information Science and Technology \textbf{58} (5), pp 702-09,
  2007.

\bibitem{Kornai1999}
  A. Kornai,
  \emph{Zipf's law outside the middle range}, in J. Rogers, \emph{Proceedings of the Sixth Meeting on Mathematics of Language}, pp 347-56.
  University of Central Florida,
  1999.

\bibitem{Milicka2009}
  J. Milička,
  \emph{Type-token \& Hapax-token Relation: A Combinatorial Model}.
  Glottotheory. International Journal of Theoretical Linguistics \textbf{1} (2), pp 99-110,
  2009.

\bibitem{Shang2010}
  Ming-Sheng Shang, Linyuan Lu, Yi-Cheng Zhang, Tao Zhou,
  \emph{Empirical analysis of web-based user-object bipartite networks}.
  A Letters Journal Exploring the Frontiers of Physics, EPL \textbf{90}, 48006,
  2010.

\bibitem{LindGonzalezHerrmann2005}
  P.G. Lind, M.C. González, H.J. Herrmann,
  \emph{Cycles and clustering in bipartite networks}.
  Physical Review E \textbf{72}, 056127,
  2005.

\bibitem{Jaccard1901}
  P. Jaccard,
  \emph{Étude comparative de la distribution florale dans une portion des Alpes et des Jura}.
  Bulletin de la Société Vaudoise des Sciences Naturelles \textbf{37}, pp 547-79,
  1901.

\bibitem{MasoudiSchreiberKashani2012}
  A. Masoudi-Nejad, F. Schreiber, Z.R. Kashani
  \emph{Building Blocks of Biological Networks: A Review on Major Network Motif Discovery Algorithms}.
  IET Systems Biology \textbf{6} (5), pp 164-74,
  2012.

\bibitem{MiloAlon2002}
  R. Milo, U. Alon, \emph{et al.},
  \emph{Network Motifs: Simple Building Blocks of Complex Networks}.
  Science \textbf{298} (5594), pp 824-27,
  2002.

\bibitem{ManganZaslaverAlon2003}
  S. Mangan, A. Zaslaver, U. Alon,
  \emph{The coherent feedforward loop serves as a sign-sensitive delay element in transcription networks}.
  Journal of Molecular Biology \textbf{334} (2), pp 197-204,
  2003.

\bibitem{Eichenberger2004}
  P. Eichenberger, \emph{et al.},
  \emph{The program of gene transcription for a single differentiating cell type during sporulation in Bacillus subtilis}.
  PLoS Biol. \textbf{2} (10), pp 1664-83,
  2004.

\bibitem{Lee2002}
  T.I. Lee, \emph{et al.},
  \emph{Transcriptional regulatory networks in Saccharomyces cerevisiae}.
  Science \textbf{298} (5594), pp 799-804,
  2002.

\bibitem{Saddic2006}
  L.A. Saddic, \emph{et al.},
  \emph{The LEAFY target LMI1 is a meristem identity regulator and acts together with LEAFY to regulate expression of CAULIFLOWER}.
  Development \textbf{133} (9), pp 1673-82,
  2006.

\bibitem{Boyer2005}
  L.A. Boyer, \emph{et al.},
  \emph{Core transcriptional regulatory circuitry in human embryonic stem cells}.
  Cell \textbf{122} (6), pp 947-56,
  2005.

\bibitem{Alon2007}
  U. Alon,
  \emph{Network motifs: theory and experimental approaches}.
  Nature Reviews Genetics \textbf{8} (6), pp 450-61,
  2007.
  
\bibitem{FischerArnold1990}
  E. Fischer, S.J. Arnold,
  \emph{More than a Labor of Love: Gender Roles and Christmas Gift Shopping}
  Journal of Consumer Research, \textbf{17} (3), pp 333-45,
  1990.

\bibitem{ArnoldReynolds2003}
  M.J. Arnold, K.E. Reynolds,
  \emph{Hedonic shopping motivations}.
  Journal of Retailing \textbf{79} (2), pp 77-95,
  2003.

\bibitem{GebertPries2012}
  Steffen Gebert, Rastin Pries, Daniel Schlosser, Klaus Heck,
  \emph{Internet Access Traffic Measurement and Analysis}.
  TMA'12 Proceedings of the 4th international conference on Traffic Monitoring and Analysis,
  Springer-Verlag Berlin, Heidelberg,
  2012.

\bibitem{Carskadon2002}
  M.A. Carskadon
  \emph{Adolescent Sleep Patterns: Biological, Social, and Psychological Influences.}.
  Cambridge University Press,
  2002.

\bibitem{ToivonenMahlerZhou2010}
  H. Toivonen, S. Mahler, F. Zhou,
  \emph{A Framework for Path-Oriented Network Simplification}.
  in P.R. Cohen, N.M. Adams, M.R. Berthold,
  IDA 2010. LNCS, \textbf{6065}, pp 220–31,
  Springer, Heidelberg,
  2010.

\bibitem{ZhouMahlerToivonen2012}
  F. Zhou, S. Mahler, H. Toivonen,
  \emph{Simplification of Networks by Edge Pruning}.
  Bisociative Knowledge Discovery,
  Lecture Notes in Computer Science \textbf{7250}, pp 179-98,
  2012.

\end{thebibliography}
